---
title: "week-8-hw-eeb174.Rmd"
output: html_document
---

# 6

```{r}
library(ggplot2)


canids <- read.csv("/home/eeb177-student/Desktop/eeb-177/hw/canids_summary.csv", header =F, as.is = T)

names(canids) <- c("genus", "species", "minage", "maxage")
head(canids)
```

```{r}
canid_occ <- ggplot(canids, aes( species, ymin = maxage, ymax=minage, colour = genus))
canid_occ <- canid_occ + geom_linerange()
canid_occ <- canid_occ + theme(legend.position="none") #gets rid of legend
canid_occ <- canid_occ + coord_flip() #flips the coordinats horizontally
canid_occ <- canid_occ +  theme(axis.text.y = element_text(size=3)) # resizes text
canid_occ <- canid_occ + theme(axis.ticks.y=element_blank()) # gets rid of tick marks on y axis
canid_occ <- canid_occ + scale_y_continuous(limits=c(0, 40), expand = c(0, 0), breaks=c(0, 10, 20, 30, 40)) # gets rid of space on right and left of the graph to make the graph itself larger
canid_occ <- canid_occ + labs(title = "Canid Fossil Occurrences", x = "Species", y = "Ma ago") + theme(plot.title = element_text(hjust = 0.5, size=22, face = "bold"), axis.title =element_text(size=20))  # adds titles and labels to the graph
canid_occ

```

```{r}
ggsave(filename = "canid_fossil_occurrences.pdf", plot = canid_occ)

```
# 7

What taxonomic group is being displayed?
```{r}
print("Canids")
```
How many unique fossil species are contained within it?
```{r}
cat("number of unique species: ", nrow(canids))

```
How many unique fossil genera?
```{r}
genera <- unique(canids$genus)
cat("number of unique fossil genera: ", length(genera))
```
What species has the longest duration?
```{r}
canids$longest_duration <- canids$maxage - canids$minage

taxa_longest_duration<- which.max(canids$longest_duration) 

cat(canids[taxa_longest_duration,2], canids[taxa_longest_duration,5], "million years ago")
```

# 6. mustelidae
```{r}
mustelidae <- read.csv("//home/eeb177-student/Desktop/eeb-177/hw/mustelidae_summary.csv", header =F, as.is = T)

names(mustelidae) <- c("genus", "species", "minage", "maxage")
head(mustelidae)
```
```{r}
mustelidae_occ <- ggplot(mustelidae, aes( species, ymin = maxage, ymax=minage, colour = genus))
mustelidae_occ <- mustelidae_occ + geom_linerange()
mustelidae_occ <- mustelidae_occ + theme(legend.position="none") #gets rid of legend
mustelidae_occ <- mustelidae_occ + coord_flip() #flips the coordinats horizontally
mustelidae_occ <- mustelidae_occ +  theme(axis.text.y = element_text(size=3)) # resizes text
mustelidae_occ <- mustelidae_occ + theme(axis.ticks.y=element_blank()) # gets rid of tick marks on y axis
mustelidae_occ <- mustelidae_occ + scale_y_continuous(limits=c(0, 30), expand = c(0, 0), breaks=c(0, 5, 10, 15, 20, 25, 30)) # gets rid of space on right and left of the graph to make the graph itself larger
mustelidae_occ <- mustelidae_occ + labs(title = "Canid Fossil Occurrences", x = "Species", y = "Ma ago") + theme(plot.title = element_text(hjust = 0.5, size=22, face = "bold"), axis.title =element_text(size=20))  # adds titles and labels to the graph
mustelidae_occ
```

```{r}
ggsave(filename = "mustelidae_fossil_occurrences.pdf", plot = mustelidae_occ)
```

# 7. mustelidae

1. What taxonomic group is being displayed?
```{r}
print("Mustelidae")
```

2. How many unique fossil species are contained within it?
```{r}
cat("number of unique species: ", nrow(mustelidae))
```

3. How many unique fossil genera?
```{r}
genera <- unique(mustelidae$genus)
cat("number of unique fossil genera: ", length(genera))
```

4. Which species has the longest duration?
```{r}
mustelidae$longest_duration <- mustelidae$maxage - mustelidae$minage

must_longest_duration<- which.max(mustelidae$longest_duration) 

cat(mustelidae[must_longest_duration,2], canids[must_longest_duration,5], "million years ago")
```
# Part 2: Data Frames

```{r}
cats <- read.csv(file = "/home/eeb177-student/Desktop/eeb-177/class-assignments/data/feline-data.csv")
age <- c(4,5,8)
cats <- cbind(cats, age)
```

```{r}
levels(cats$coat)
levels(cats$coat) <- c(levels(cats$coat), 'tortoiseshell')
cats <- rbind(cats, list("tortoiseshell", 3.3, TRUE, 9))
cats

```
```{r}
str(cats)

```
```{r}
cats$coat <- as.character(cats$coat)
str(cats)

```
```{r}
cats[-4,]
```
```{r}
cats <- na.omit(cats)

```
```{r}
cats <- rbind(cats, cats)
cats
rownames(cats) <- NULL
cats
```
# Challenge 1

```{r}
df <- data.frame(first = c('Grace'),
                 last = c('Hopper'),
                 lucky_number = c(0),
                 stringsAsFactors = FALSE)
df <- rbind(df, list('Marie', 'Curie', 238) )
df <- cbind(df, coffeetime = c(TRUE,TRUE))
df
```
```{r}
gapminder <- read.csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder-FiveYearData.csv")
```
```{r}
str(gapminder)
```
```{r}
typeof(gapminder$year)
```
```{r}
typeof(gapminder$country)
```
```{r}
str(gapminder$country)
```
```{r}
length(gapminder)
```
```{r}
typeof(gapminder)
```
```{r}
dim(gapminder)
```
```{r}
colnames(gapminder)

```
```{r}
head(gapminder)
```

# Challenge 2
```{r}
gapminder <- read.csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder-FiveYearData.csv")
#source(file = "~/scripts/load-gapminder.R")
```
# Challenge 3
gapminder = data frame. columns with 1704 rows, 6 columns. 
Country and continent = factors
Year = integer vector
Pop = numeric vector
lifeExp = numeric vector
gdpPercap  = numeric vector

#2. Subsetting Data

```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
x
```
```{r}
x[1]
x[4]
x[c(1,3)]
x[1:4]
```
```{r}
x[c(1,1,3)]
x[6]
x[0]
```
```{r}
x[-2]
x[c(-1, -5)]
```
```{r}
x <- x[-4]
x
```
# Challenge 1
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)

x[2:4]
x[-c(1,5)]
x[c("b", "c", "d")]
x[c(2,3,4)]
```
```{r}
x[c("a", "c")]
```
```{r}
x[-which(names(x) == "a")]
names(x) == "a"
which(names(x) == "a")
x[-which(names(x) %in% c("a", "c"))]
```
# Challenge 2
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)
```
```{r}
x[-which(names(x) == "g")]
```

if there are no TRUE values passed to which, an empty vector is returned integer() is an empty vector

```{r}
names(x) <- c('a', 'a', 'a')
x['a']
x[which(names(x) == 'a')]
```
```{r}
names(x) == c('a', 'c', 'e')
```
```{r}
x[c(TRUE, TRUE, FALSE, FALSE)]
x[c(TRUE, FALSE)]
x[x > 7]
```
# Challenge 3
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)

x_subset <- x[x<7 & x>4]
print(x_subset)
```
```{r}
f <- factor(c("a", "a", "b", "c", "c", "d"))
f[f == "a"]
```
```{r}
f[f %in% c("b", "c")]
f[1:3]
f[-3]
```
```{r}
set.seed(1)
m <- matrix(rnorm(6*4), ncol=4, nrow=6)
m[3:4, c(3,1)]
```
```{r}
m[, c(3,4)]
```
```{r}
m[3,]
m[3, , drop=FALSE]
```
```{r}
m[5]
```
```{r}
matrix(1:6, nrow=2, ncol=3)
```
```{r}
matrix(1:6, nrow=2, ncol=3, byrow=TRUE)
```
# Challenge 4
```{r}
m <- matrix(1:18, nrow=3, ncol=6)
print(m)
```
m[2,c(4,5)] will extract the values 11 and 14.
```{r}
xlist <- list(a = "Software Carpentry", b = 1:10, data = head(iris))
xlist[1]
```
```{r}
xlist[1:2]
```
```{r}
xlist[[1]]
```
```{r}
xlist[["a"]]
```
```{r}
xlist$data
```
# Challenge 5
```{r}
xlist <- list(a = "Software Carpentry", b = 1:10, data = head(iris))
xlist[[2]][2]
```
# Challenge 6
```{r}
mod <- aov(pop ~ lifeExp, data=gapminder)
attributes(mod)
mod$df.residual
```
```{r}
head(gapminder[3])
```
```{r}
head(gapminder[["lifeExp"]])
```
```{r}
head(gapminder$year)
```
```{r}
gapminder[1:3,]
```
```{r}
gapminder[3,]
```
# Challenge 7
```{r}
gapminder[gapminder$year == 1957,]
```
```{r}
gapminder[,-c(1:4)]
```
```{r}
gapminder[gapminder$lifeExp > 80,]
```
```{r}
gapminder[1, c(4, 5)]
```
```{r}
gapminder[gapminder$year == 2002 | gapminder$year == 2007,]
gapminder[gapminder$year %in% c(2002, 2007),]
```
# Challenge 8
gapminder is a data frame so it needs to be subsetted on 2 dimensions. gapminder[1:20, ] subsets the data to give the first 20 rows and all columns.
```{r}
gapminder_small <- gapminder[c(1:9, 19:23),]
```
# 3. Creating Publication Quality Graphics
```{r}
library("ggplot2")
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point()
```
# Challenge 1
```{r}
ggplot(data = gapminder, aes(x = year, y = lifeExp)) + geom_point()
```
# Challenge 2
```{r}
ggplot(data = gapminder, aes(x = year, y = lifeExp, color=continent)) +
  geom_point()
```
```{r}
ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country, color=continent)) +
  geom_line()
```
```{r}
ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country, color=continent)) +
  geom_line() + geom_point()
```
```{r}
ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country)) +
  geom_line(aes(color=continent)) + geom_point()
```
# Challenge 3
The lines get drawn over the points.
```{r}
ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country)) +
 geom_point() + geom_line(aes(color=continent))
```
```{r}
ggplot(data = gapminder, aes(x=year, y=lifeExp, by=country)) +
 geom_point() + geom_line(aes(color=continent))
```
```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color=continent)) +
  geom_point()
```
```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point(alpha = 0.5) + scale_x_log10()
```
```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() + scale_x_log10() + geom_smooth(method="lm")
```
```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
  geom_point() + scale_x_log10() + geom_smooth(method="lm", size=1.5)
```
# Challenge 4a
```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) +
 geom_point(size=3, color="orange") + scale_x_log10() +
 geom_smooth(method="lm", size=1.5)
```
# Challenge 4b
```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) +
geom_point(size=3, pch=17) + scale_x_log10() +
geom_smooth(method="lm", size=1.5)
```
```{r}
starts.with <- substr(gapminder$country, start = 1, stop = 1)
az.countries <- gapminder[starts.with %in% c("A", "Z"), ]
ggplot(data = az.countries, aes(x = year, y = lifeExp, color=continent)) +
  geom_line() + facet_wrap( ~ country)
```
```{r}
ggplot(data = az.countries, aes(x = year, y = lifeExp, color=continent)) +
  geom_line() + facet_wrap( ~ country) +
  xlab("Year") + ylab("Life expectancy") + ggtitle("Figure 1") +
  scale_colour_discrete(name="Continent") +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
```
# Challenge 5
```{r}
ggplot(data = gapminder, aes(x = gdpPercap, fill=continent)) +
 geom_density(alpha=0.6) + facet_wrap( ~ year) + scale_x_log10()
```




















